---
title: "hw2 q17"
author: "313707025 jebuhdah"
date: "2025-03-09"
output: 
  pdf_document:
    latex_engine: xelatex

---
```{r load_library, echo=FALSE, message=FALSE, warning=FALSE }
library(ggplot2)
library(readr)

# load data
data <- read_csv("collegetown.csv")
```
'######################################################################################################
qn17:
a:
```{r qa, echo=FALSE, message=FALSE, warning=FALSE }
# plot PRICE vs SQFT
ggplot(data, aes(x = sqft, y = price)) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(title = "Scatter Plot of House Price vs House Size",
       x = "House Size (SQFT)",
       y = "House Price (PRICE)") +
  theme_minimal()
```
the trend is upward... implying larger houses tends to have higher prices, not a surprise

```{r qb, echo=FALSE, message=FALSE, warning=FALSE }
# fit the linear regression model
linear_model <- lm(price ~ sqft, data = data)

# the model
summary(linear_model)
##########################################
ggplot(data, aes(x = sqft, y = price)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Linear Regression Line",
       x = "House Size (SQFT)",
       y = "House Price (PRICE)") +
  theme_minimal()

# Fit quadratic regression model
quadratic_model <- lm(price ~ I(sqft^2), data = data)
```
b:as per beta by above summary, the linear regression model price=B1+B2sqft+e: price=-115.42+13.40sqft. which implies that a house the size of 0sqft... will start off at -115423.6 dollars, and when everything is constant, 100 sqft increase in house size will cause the price to raise by 13402 dollars.

```{r qc, echo=FALSE, message=FALSE, warning=FALSE }
# the model
summary(quadratic_model)

# plot the quadratic curve
ggplot(data, aes(x = sqft, y = price)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", formula = y ~ I(x^2), color = "red", se = FALSE) +
  labs(title = "Fitted Quadratic Regression Curve",
       x = "House Size (SQFT)",
       y = "House Price (PRICE)") +
  theme_minimal()
```
c:for price=a1+a2(sqft^2)+e, price=93.57+0.185(sqft^2)
the marginal effect for 20k sqft is 2(a2)sqft = 2(0.185)20^2 = 7.4 (about)

```{r qd, echo=FALSE, message=FALSE, warning=FALSE }
# plot the quadratic curve
ggplot(data, aes(x = sqft, y = price)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", formula = y ~ I(x^2), color = "red", se = FALSE) +
  labs(title = "Fitted Quadratic Regression Curve",
       x = "House Size (SQFT)",
       y = "House Price (PRICE)") +
  theme_minimal()
```
d: as shown above
e: elasticity = (∂price/∂sqft)(sqft/price) = 2(0.185)(20)(20)/(93.57+0.185(20^2))=0.8832

```{r qf, echo=FALSE, message=FALSE, warning=FALSE }
# residuals for linear model
data$linear_residuals <- residuals(linear_model)

# residuals for quadratic model
data$quadratic_residuals <- residuals(quadratic_model)

# plot residuals for linear 
ggplot(data, aes(x = sqft, y = linear_residuals)) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(title = "Residuals from Linear Model",
       x = "House Size (SQFT)",
       y = "Residuals") +
  theme_minimal()

# plot residuals for quadratic 
ggplot(data, aes(x = sqft, y = quadratic_residuals)) +
  geom_point(color = "red", alpha = 0.6) +
  labs(title = "Residuals from Quadratic Model",
       x = "House Size (SQFT)",
       y = "Residuals") +
  theme_minimal()
```
f:it seems like in the plots the residual variation increases with the x axis for both the models which likely violates homoskedasticity.

```{r qg, echo=FALSE, message=FALSE, warning=FALSE }
# SSE linear model
linear_SSE <- sum(residuals(linear_model)^2)

# SSE quadratic model
quadratic_SSE <- sum(residuals(quadratic_model)^2)

# show me
cat("Linear Model SSE:", linear_SSE, "\n")
cat("Quadratic Model SSE:", quadratic_SSE, "\n")
```
g:the modle from c has a lower sse, and this implies the data is closer to the fitted line of the quadratic model in this case.