---
title: "HW0324"
author: "Yung-Jung Cheng"
date: "2025-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(POE5Rdata)

```

# Q1
Let $K=2$, show that $(b_1, b_2)$ in p.29 of slide in Ch 5 reduces to the formual of $(b_1,b_2)$ in (2.7)-(2.8)

### Ans
##### Step 1: Model and Objective Function

Consider the simple linear regression model:

$$
y_i = b_1 + b_2 x_i + e_i, \quad i = 1, 2, \dots, N
$$

The least squares method minimizes the sum of squared errors (SSE):

$$
SSE(b_1, b_2) = \sum_{i=1}^N (y_i - b_1 - b_2 x_i)^2
$$

---

##### Step 2: Take Derivatives with Respect to $b_1$ and $b_2$

**Partial derivative w.r.t. $b_1$:**

$$
\frac{\partial SSE}{\partial b_1} = -2 \sum (y_i - b_1 - b_2 x_i) = 0
$$

Solving:

$$
\sum y_i = N b_1 + b_2 \sum x_i \quad \Rightarrow \quad b_1 = \overline{y} - b_2 \overline{x} 
$$

**Partial derivative w.r.t. $b_2$:**

$$
\frac{\partial SSE}{\partial b_2} = -2 \sum x_i (y_i - b_1 - b_2 x_i) = 0
$$

Substitute $b_1 = \overline{y} - b_2 \overline{x}$:

$$
\sum x_i (y_i - \overline{y} + b_2 \overline{x} - b_2 x_i) = 0
$$

Simplifying:

$$
\sum (x_i - \overline{x})(y_i - \overline{y}) = b_2 \sum (x_i - \overline{x})^2
$$

Thus:

$$
b_2 = \frac{\sum (x_i - \overline{x})(y_i - \overline{y})}{\sum (x_i - \overline{x})^2}
$$

---

#####  Conclusion from LSE Derivation:

We obtain the closed-form solution:

$$
b_2 = \frac{\sum (x_i - \overline{x})(y_i - \overline{y})}{\sum (x_i - \overline{x})^2}, \quad b_1 = \overline{y} - b_2 \overline{x}
$$

---

#### Verify Equivalence with Matrix Formula $b = (X'X)^{-1} X'Y$

Define:

$$
X = \begin{bmatrix}
1 & x_1 \\
1 & x_2 \\
\vdots & \vdots \\
1 & x_N
\end{bmatrix}, \quad
Y = \begin{bmatrix}
y_1 \\
y_2 \\
\vdots \\
y_N
\end{bmatrix}
$$

Compute:

$$
X'X = \begin{bmatrix}
N & \sum x_i \\
\sum x_i & \sum x_i^2
\end{bmatrix}, \quad
X'Y = \begin{bmatrix}
\sum y_i \\
\sum x_i y_i
\end{bmatrix}
$$

Then:

$$
(X'X)^{-1} = \frac{1}{N \sum x_i^2 - (\sum x_i)^2}
\begin{bmatrix}
\sum x_i^2 & -\sum x_i \\
-\sum x_i & N
\end{bmatrix}
$$

So:

$$
b = (X'X)^{-1} X'Y =
\frac{1}{D}
\begin{bmatrix}
\sum x_i^2 \sum y_i - \sum x_i \sum x_i y_i \\
N \sum x_i y_i - \sum x_i \sum y_i
\end{bmatrix}
$$

Where $D = N \sum x_i^2 - (\sum x_i)^2$.

---

#####  Final Transformation

Using the identity:

$$
\sum (x_i - \overline{x})(y_i - \overline{y}) = \sum x_i y_i - N \overline{x} \, \overline{y}
$$

We can show:

$$
b_2 = \frac{\sum (x_i - \overline{x})(y_i - \overline{y})}{\sum (x_i - \overline{x})^2}, \quad b_1 = \overline{y} - b_2 \overline{x}
$$

---

##### Conclusion

We have shown that the scalar-form least squares solution $(b_1, b_2)$ derived from minimizing SSE is equivalent to the matrix-form solution:

$$
b = (X'X)^{-1} X'Y
$$

for the case $K = 2$

---

# Q2

Let $K=2$, show that $cov(b_1,b_2)$ in p.30 of slide in Ch 5 reduces to the formula of in (2.14)-(2.16)

### Ans


#### Step 1: General variance formula for OLS estimators

Given the linear model:

$$
Y = X\beta + \varepsilon, \quad \text{with } \operatorname{Var}(\varepsilon) = \sigma^2 I,
$$

the least squares estimator is:

$$
\hat{\beta} = (X'X)^{-1} X'Y
$$

and its variance-covariance matrix is:

$$
\operatorname{Var}(\hat{\beta}) = \sigma^2 (X'X)^{-1}
$$

---

####  Step 2: Compute $(X'X)^{-1}$ for simple linear regression ($K = 2$)

Let:

$$
X = \begin{bmatrix}
1 & x_1 \\
1 & x_2 \\
\vdots & \vdots \\
1 & x_N
\end{bmatrix}
$$

Then:

$$
X'X = \begin{bmatrix}
N & \sum x_i \\
\sum x_i & \sum x_i^2
\end{bmatrix}
$$

Using the $2 \times 2$ inverse formula:

$$
(X'X)^{-1} = \frac{1}{D}
\begin{bmatrix}
\sum x_i^2 & -\sum x_i \\
-\sum x_i & N
\end{bmatrix}, \quad \text{where } D = N \sum x_i^2 - (\sum x_i)^2
$$

---

####  Step 3: Multiply by $\sigma^2$ to get the covariance matrix

$$
\operatorname{Var}(b) = \sigma^2 (X'X)^{-1} =
\frac{\sigma^2}{D}
\begin{bmatrix}
\sum x_i^2 & -\sum x_i \\
-\sum x_i & N
\end{bmatrix}
$$

Extract components:

- $\operatorname{Var}(b_1) = \frac{\sigma^2 \sum x_i^2}{D}$
- $\operatorname{Var}(b_2) = \frac{\sigma^2 N}{D}$
- $\operatorname{Cov}(b_1, b_2) = -\frac{\sigma^2 \sum x_i}{D}$

---

####  Step 4: Express $D$ using centered data

Define:

$$
\overline{x} = \frac{1}{N} \sum x_i, \quad S_{xx} = \sum (x_i - \overline{x})^2
$$

Then:

$$
D = N \sum x_i^2 - (\sum x_i)^2 = N S_{xx}
$$

---

#### Step 5: Express variances in terms of $S_{xx}$ and $\overline{x}$

Use identity:

$$
\sum x_i^2 = S_{xx} + N \overline{x}^2
$$

Substitute into each term:

- $\operatorname{Var}(b_1) = \frac{\sigma^2 (S_{xx} + N \overline{x}^2)}{N S_{xx}} = \frac{\sigma^2}{N} + \frac{\sigma^2 \overline{x}^2}{S_{xx}}$
- $\operatorname{Var}(b_2) = \frac{\sigma^2}{S_{xx}}$
- $\operatorname{Cov}(b_1, b_2) = -\frac{\sigma^2 \overline{x}}{S_{xx}}$

---

#### Final Result (as in page 30)

$$
\operatorname{Var}(b_1) = \frac{\sigma^2}{N} + \frac{\sigma^2 \overline{x}^2}{\sum (x_i - \overline{x})^2}
$$

$$
\operatorname{Var}(b_2) = \frac{\sigma^2}{\sum (x_i - \overline{x})^2}
$$

$$
\operatorname{Cov}(b_1, b_2) = - \frac{\sigma^2 \overline{x}}{\sum (x_i - \overline{x})^2}
$$

These match exactly with the expressions on slide page 30.

---

# Q03

Consider the following model that relates the percentage of a household’s budget spent on alcohol $WALC$ to total expenditure $TOTEXP$, age of the household head $AGE$, and the number of children in the household $NK$:

$$
WALC = \beta_1 + \beta_2 \ln(TOTEXP) + \beta_3 NK + \beta_4 AGE + e
$$

This model was estimated using 1200 observations from London. An incomplete version of this output is provided in Table 5.6.

---

### TABLE 5.6 Output for Exercise 5.3

**Dependent Variable:** WALC  
**Included observations:** 1200

| Variable        | Coefficient | Std. Error | t-Statistic | Prob. |
|-----------------|-------------|------------|-------------|-------|
| C               | 1.4515      | 2.2019     | ?           | 0.5099|
| ln(TOTEXP)      | 2.7648      | ?          | 5.7103      | 0.0000|
| NK              | ?           | 0.3695     | -3.9376     | 0.0001|
| AGE             | -0.1503     | 0.0235     | -6.4019     | 0.0000|

- R-squared: ?  
- S.E. of regression: 6.39547  
- Sum squared resid: 46221.62  
- Mean dependent var: 6.19434  
- S.D. dependent var: 6.39547  

---


## Q3(a) 
Fill in the following blank spaces that appear in this table.  
  i. The $t$-statistic for $\beta_1$  
  ii. The standard error for $\beta_2$  
  iii. The estimate $\beta_3$  
  iv. $R^2$  
  v. $\hat{\sigma}$  

### Ans
  i. The $t$-statistic for $\beta_1$  $\boxed{0.6593}$
$$
t = \frac{\hat{\beta}_1}{\text{SE}(\hat{\beta}_1)} = \frac{1.4515}{2.2019} ≈ 0.6593
$$


  ii. The standard error for $\beta_2$  $\boxed{0.4840}$
$$
\text{SE} = \frac{\hat{\beta}_2}{t} = \frac{2.7648}{5.7103} ≈ 0.4840
$$
  iii. The estimate $\beta_3$  $\boxed{-1.4548}$

$$
\hat{\beta}_3 = t \cdot \text{SE} = (-3.9376)(0.3695) ≈ -1.4548
$$
  iv. $R^2$  $\boxed{0.0588}$
$$
R^2 = 1 - \frac{SSR}{TSS}
$$

  - SSR = 46221.62
  - TSS = $n \cdot \text{S.D.}^2 = 1200 \cdot (6.39547)^2 ≈ 49107.84$

$$
R^2 = 1 - \frac{46221.62}{49107.84} ≈ 1 - 0.9412 = 0.0588
$$
  v. $\hat{\sigma}$  $\boxed{6.2180}$
$$
\hat{\sigma} = \sqrt{\frac{SSR}{n - K - 1}} = \sqrt{\frac{46221.62}{1196}} = \sqrt{38.6529} ≈ 6.2180
$$




## Q3(b) 
Interpret each of the estimates $\beta_2$, $\beta_3$, and $\beta_4$.  

### Ans
| Variable     | Marginal Effect Formula                    | Economic Interpretation                                                      |
|--------------|---------------------------------------------|-------------------------------------------------------------------------------|
| $\hat{\beta}_2$  | $\partial WALC / \partial \ln(TOTEXP)$      | A 1% increase in total expenditure increases alcohol share by 2.7648 percentage points. |
| $\hat{\beta}_3$  | $\partial WALC / \partial NK$               | One additional child reduces alcohol share by 1.4548 percentage points.       |
| $\hat{\beta}_4$  | $\partial WALC / \partial AGE$              | One more year of age reduces alcohol share by 0.1503 percentage points.       |

## Q3(c)
Compute a 95% interval estimate for $\beta_4$. What does this interval tell you?  

### Ans
We are given:
- $\hat{\beta}_4 = -0.1503$
- Standard error $SE(\hat{\beta}_4) = 0.0235$
- $z_{0.025} = 1.96$ for a 95% confidence level

**Step 1**: Compute the margin of error  
$ME = 1.96 \times 0.0235 = 0.0461$

**Step 2**: Construct the confidence interval  
Lower bound: $-0.1503 - 0.0461 = -0.1964$  
Upper bound: $-0.1503 + 0.0461 = -0.1042$

**Conclusion**:  
The 95% confidence interval for $\beta_4$ is $(-0.1964, -0.1042)$.  
We are 95% confident that the effect of age on alcohol expenditure share lies within this interval.  
Since the entire interval is negative, this implies that age has a statistically significant negative effect.


## Q3(d)
Are each of the coefficient estimates significant at a 5% level? Why?  

### Ans

To determine whether each coefficient is statistically significant at the 5% level, we compare their p-values with 0.05.

| Coefficient | Estimate  | t-Statistic | p-value | Significant at 5%? |
|-------------|-----------|-------------|---------|---------------------|
| $\beta_1$   | 1.4515    | 0.6593      | 0.5099  | No                  |
| $\beta_2$   | 2.7648    | 5.7103      | 0.0000  | Yes                 |
| $\beta_3$   | -1.4548   | -3.9376     | 0.0001  | Yes                 |
| $\beta_4$   | -0.1503   | -6.4019     | 0.0000  | Yes                 |

**Conclusion**:  
At the 5% significance level, $\beta_2$, $\beta_3$, and $\beta_4$ are statistically significant, while $\beta_1$ is not.


## Q3(e)
Test the hypothesis that the addition of an extra child decreases the mean budget share of alcohol by 2 percentage points against the alternative that the decrease is not equal to 2 percentage points.  Use a 5% significance level.

### Ans

**Null hypothesis**: $H_0: \beta_3 = -2$  
**Alternative hypothesis**: $H_1: \beta_3 \neq -2$  
**Significance level**: $\alpha = 0.05$

We are given:
- $\hat{\beta}_3 = -1.4548$
- $SE(\hat{\beta}_3) = 0.3695$

**Step 1**: Compute the test statistic  
$t = \frac{-1.4548 - (-2)}{0.3695} = \frac{0.5452}{0.3695} ≈ 1.4755$

**Step 2**: Compare with critical value  
The critical value at 5% significance level (two-tailed) is $1.96$.  
Since $|t| = 1.4755 < 1.96$, we fail to reject the null hypothesis.

**Conclusion**:  
There is not enough evidence at the 5% level to conclude that $\beta_3$ differs from -2.


---
# Q23
The file `cocaine` contains 56 observations on variables related to sales of cocaine powder in northeastern California over the period 1984–1991. The data are a subset of those used in the study: Caulkins, J. P. and R. Padman (1993), “Quantity Discounts and Quality Premia for Illicit Drugs,” *Journal of the American Statistical Association*, 88, 748–757. The variables are:
- `PRICE`: price per gram in dollars for a cocaine sale  
- `QUANT`: number of grams of cocaine in a given sale  
- `QUAL`: quality of the cocaine expressed as percentage purity  
- `TREND`: a time variable with 1984 = 1 up to 1991 = 8  
Consider the regression model:
$$
PRICE = \beta_1 + \beta_2 QUANT + \beta_3 QUAL + \beta_4 TREND + e
$$

## Q23(a)
What signs would you expect on the coefficients $\beta_2$, $\beta_3$, and $\beta_4$?

### Ans

- $\beta_2$: Expected to be **negative** due to quantity discounts — as quantity increases, price per gram decreases.
- $\beta_3$: Expected to be **positive** because higher purity should command a higher price.
- $\beta_4$: Likely **negative** if price has declined over time due to increased availability or law enforcement pressure.

## Q23(b)
Use your computer software to estimate the equation. Report the results and interpret the coefficient estimates. Have the signs turned out as you expected?

### Ans
```{r Q23b}
model <- lm(price ~ quant + qual + trend, data = cocaine)
summary(model)

```

The estimated regression equation is:

$$
\widehat{PRICE} = 90.8467 - 0.0600 \cdot QUANT + 0.1162 \cdot QUAL - 2.3546 \cdot TREND
$$

**Interpretation**:

- **QUANT**: The coefficient is -0.0600 and is statistically significant at the 1% level. This confirms the presence of quantity discounts: as quantity increases by one gram, the price per gram drops by 0.06 dollars.
- **QUAL**: The coefficient is 0.1162, which has the expected positive sign, suggesting higher purity leads to higher price. However, it is **not statistically significant** (p = 0.5700), so we cannot confidently say that purity affects price.
- **TREND**: The coefficient is -2.3546 with a p-value of 0.0954, which is marginally significant at the 10% level. This implies a possible downward trend in price over time.

**Conclusion**:  
The signs of all coefficients match the expectations from part (a), but only **QUANT** is strongly significant. **QUAL** is not significant, and **TREND** is weakly significant at the 10% level.



## Q23(c)
What proportion of variation in cocaine price is explained jointly by variation in quantity, quality, and time?

### Ans

The model's R-squared is 0.5097, meaning that approximately **50.97%** of the variation in cocaine prices is explained by the variation in quantity, quality, and time (trend).

**Interpretation**:  
This suggests that about half of the price variation can be accounted for by these three variables. The remaining variation may be due to other factors not included in the model, such as transaction location, seller identity, buyer type, or law enforcement pressure.



## Q23(d)
It is claimed that the greater the number of sales, the higher the risk of getting caught. Thus, sellers are willing to accept a lower price if they can make sales in larger quantities.  
Set up $H_0$ and $H_1$ that would be appropriate to test this hypothesis. Carry out the hypothesis test.

### Ans

**Claim**: Sellers are willing to accept a lower price when selling in larger quantities.

We test:

- Null hypothesis: $H_0: \beta_2 = 0$
- Alternative hypothesis: $H_1: \beta_2 < 0$
- Significance level: $\alpha = 0.05$

From the regression output:
- Estimate: $\hat{\beta}_2 = -0.05997$
- Standard error: $SE = 0.01018$
- Test statistic: $t = -5.892$
- p-value: $2.85 \times 10^{-7}$

**Conclusion**:  
Since the p-value is far below 0.05 and the t-statistic is less than the critical value, we reject the null hypothesis.  
There is strong evidence to support the existence of a quantity discount in cocaine pricing.




## Q23(e)
Test the hypothesis that the quality of cocaine has no influence on expected price against the alternative that a premium is paid for better-quality cocaine.

### Ans

We test whether better-quality cocaine commands a higher price.

- Null hypothesis: $H_0: \beta_3 = 0$  
- Alternative hypothesis: $H_1: \beta_3 > 0$  
- Significance level: $\alpha = 0.05$

From the regression output:
- Estimate: $\hat{\beta}_3 = 0.1162$
- Standard error: $SE = 0.2033$
- Test statistic: $t = 0.572$
- Critical value: $t_{0.05, 52} \approx 1.675$
- p-value = 0.5700

**Conclusion**:  
We fail to reject the null hypothesis.  
There is no statistically significant evidence that a premium is paid for better-quality cocaine.




## Q23(f)
What is the average annual change in the cocaine price? Can you suggest why price might be changing in this direction?

### Ans

From the regression results, the coefficient on the TREND variable is:

$$
\hat{\beta}_4 = -2.3546
$$

This indicates that the average cocaine price per gram **decreased by approximately \$2.35 per year** from 1984 to 1991.

**Possible Reasons**:
- Increased supply due to mass production and export from South America (e.g., Colombia)
- Competition among dealers driving prices down
- Law enforcement unable to constrain supply effectively
- Market saturation and commoditization of cocaine
- Economic or social factors reducing willingness to pay

**Conclusion**:  
There was a significant downward trend in cocaine prices over this period, likely driven by supply-side factors and increased market competition.





















