---
title: "HW0303 - CH2 Questions 17, 25, 28"
author: "鄭凱鴻"
date: "2025-03-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **CH2 Question 17**

## **(a) Scatter Plot of House Price vs. House Size**

```{r}
library(ggplot2)  # 載入 ggplot2 繪圖函式庫
load("collegetown.rdata")  # 載入數據

# 繪製房屋價格與房屋大小的散點圖
ggplot(collegetown, aes(x = sqft, y = price)) +
    geom_point(color = "blue", alpha = 0.6) +  # 添加散點，設置藍色，透明度 0.6
    labs(title = "House Price vs House Size",
         x = "House Size (hundreds of square feet)",
         y = "Price (thousands of dollars)") +
    theme_minimal()  # 設置簡潔主題
```

## **(b) Linear Regression Model and Fitted Line**

```{r}
# 執行線性回歸
model <- lm(price ~ sqft, data = collegetown)
summary(model)  # 顯示回歸結果

# 繪製回歸線
ggplot(collegetown, aes(x = sqft, y = price)) +
    geom_point(color = "blue", alpha = 0.6) +  # 散點圖
    geom_smooth(method = "lm", color = "red", se = FALSE) +  # 添加線性回歸線
    labs(title = "Linear Regression: PRICE vs. SQFT",
         x = "House Size (hundreds of square feet)",
         y = "Price (thousands of dollars)") +
    theme_minimal()
```

## **(c) Quadratic Regression Model and Marginal Effect Calculation**

```{r}
# 創建平方項
collegetown$sqft2 <- collegetown$sqft^2

# 執行二次回歸
model_quad_pure <- lm(price ~ sqft2, data = collegetown)
summary(model_quad_pure)  # 顯示回歸結果

# 計算邊際影響
alpha2 <- coef(model_quad_pure)["sqft2"]
sqft_value <- 20  # 2000 平方英尺（hundreds of square feet）
marginal_effect <- 2 * alpha2 * sqft_value

cat("當房屋面積為 2000 平方英尺時，增加 100 平方英尺對價格的影響為：", marginal_effect, "千美元\n")
```

## **(d) Quadratic Regression Curve with Tangent Line at 2000 Square Feet**

```{r}
# 提取回歸係數
alpha0 <- coef(model_quad_pure)["(Intercept)"]
alpha2 <- coef(model_quad_pure)["sqft2"]

# 設定 x0（2000 平方英尺）
x0 <- 20
y0 <- alpha0 + alpha2 * x0^2  # 計算 y0（對應價格）

# 計算切線斜率
m <- 2 * alpha2 * x0

# 定義切線函數
tangent_line <- function(x) {
    m * (x - x0) + y0
}

# 繪製二次回歸曲線與切線
ggplot(collegetown, aes(x = sqft, y = price)) +
    geom_point(color = "blue", alpha = 0.6) +
    geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw=TRUE), color = "red", se = FALSE) +
    stat_function(fun = tangent_line, color = "green", linetype = "dashed") +  # 添加切線
    geom_vline(xintercept = x0, color = "black", linetype = "dotted") +  # 標示 x0 位置
    geom_point(aes(x = x0, y = y0), color = "black", size = 3) +  # 添加切點
    labs(title = "Quadratic Regression with Tangent Line at SQFT = 2000",
         x = "House Size (hundreds of square feet)",
         y = "Price (thousands of dollars)") +
    theme_minimal()
```

## **(e) Elasticity of PRICE with Respect to SQFT at 2000 Square Feet**

```{r}
# 計算價格預測值
price_predicted <- alpha0 + alpha2 * sqft_value^2

# 計算彈性
elasticity <- (marginal_effect * sqft_value) / price_predicted
cat("當房屋面積為 2000 平方英尺時，PRICE 對 SQFT 的彈性為：", elasticity, "\n")
```

## **(f) Residual Analysis**

```{r}
# 計算線性與二次回歸的殘差
collegetown$resid_linear <- residuals(model)
collegetown$resid_quad <- residuals(model_quad_pure)

# 繪製殘差圖

# 線性回歸殘差圖
ggplot(collegetown, aes(x = sqft, y = resid_linear)) +
    geom_point(color = "blue", alpha = 0.6) +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    labs(title = "Residuals vs. SQFT (Linear Model)",
         x = "House Size (hundreds of square feet)",
         y = "Residuals") +
    theme_minimal()

# 二次回歸殘差圖

ggplot(collegetown, aes(x = sqft, y = resid_quad)) +
    geom_point(color = "blue", alpha = 0.6) +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    labs(title = "Residuals vs. SQFT (Quadratic Model)",
         x = "House Size (hundreds of square feet)",
         y = "Residuals") +
    theme_minimal()
```

## **(g) Sum of Squared Residuals (SSE) Comparison**

```{r}
# 計算 SSE（殘差平方和）
SSE_linear <- sum(residuals(model)^2)
SSE_quadratic <- sum(residuals(model_quad_pure)^2)
cat("SSE for Linear Model:", SSE_linear, "\n")
cat("SSE for Quadratic Model:", SSE_quadratic, "\n")
```

# **CH2 Question 25**

## **(a) Histogram and Summary Statistics of FOODAWAY**

```{r}
# 載入數據
load("cex5_small.rdata")

# 繪製 FOODAWAY 直方圖
library(ggplot2)
ggplot(cex5_small, aes(x = foodaway)) +
    geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
    labs(title = "Histogram of FOODAWAY",
         x = "Food Away from Home Expenditure (per person, per month)",
         y = "Frequency") +
    theme_minimal()

# 計算摘要統計量
foodaway_mean <- mean(cex5_small$foodaway, na.rm = TRUE)
foodaway_median <- median(cex5_small$foodaway, na.rm = TRUE)
foodaway_q1 <- quantile(cex5_small$foodaway, 0.25, na.rm = TRUE)
foodaway_q3 <- quantile(cex5_small$foodaway, 0.75, na.rm = TRUE)

# 顯示統計結果
cat("Mean of FOODAWAY:", foodaway_mean, "\n")
cat("Median of FOODAWAY:", foodaway_median, "\n")
cat("25th Percentile of FOODAWAY:", foodaway_q1, "\n")
cat("75th Percentile of FOODAWAY:", foodaway_q3, "\n")
```
## **(b) Mean and Median of FOODAWAY by Household Education Level**

```{r}
# 計算有高等學位者的 FOODAWAY 統計數據
subset_advanced <- cex5_small[cex5_small$advanced == 1, ]
mean_advanced <- mean(subset_advanced$foodaway, na.rm = TRUE)
median_advanced <- median(subset_advanced$foodaway, na.rm = TRUE)

# 計算有大學學位者的 FOODAWAY 統計數據
subset_college <- cex5_small[cex5_small$college == 1, ]
mean_college <- mean(subset_college$foodaway, na.rm = TRUE)
median_college <- median(subset_college$foodaway, na.rm = TRUE)

# 計算無大學或高等學位者的 FOODAWAY 統計數據
subset_no_degree <- cex5_small[cex5_small$advanced == 0 & cex5_small$college == 0, ]
mean_no_degree <- mean(subset_no_degree$foodaway, na.rm = TRUE)
median_no_degree <- median(subset_no_degree$foodaway, na.rm = TRUE)

# 輸出結果
cat("\nHousehold with Advanced Degree:\n")
cat("Mean FOODAWAY:", mean_advanced, "\n")
cat("Median FOODAWAY:", median_advanced, "\n")

cat("\nHousehold with College Degree:\n")
cat("Mean FOODAWAY:", mean_college, "\n")
cat("Median FOODAWAY:", median_college, "\n")

cat("\nHousehold with No College or Advanced Degree:\n")
cat("Mean FOODAWAY:", mean_no_degree, "\n")
cat("Median FOODAWAY:", median_no_degree, "\n")
```

## **(c) Histogram and Summary Statistics of ln(FOODAWAY)**

```{r}
# 確保 foodaway 變數存在
summary(cex5_small$foodaway)

# 檢查 FOODAWAY 是否有 0 值或 NA
cat("Number of FOODAWAY = 0:", sum(cex5_small$foodaway == 0, na.rm = TRUE), "\n")
cat("Number of missing FOODAWAY values:", sum(is.na(cex5_small$foodaway)), "\n")

# 只保留 foodaway > 0 的數據，避免 log(0) 出現 -Inf
cex5_small <- cex5_small[cex5_small$foodaway > 0, ]

# 計算 ln(FOODAWAY)
cex5_small$log_foodaway <- log(cex5_small$foodaway)

# 繪製 ln(FOODAWAY) 的直方圖
hist(cex5_small$log_foodaway, 
     main = "Histogram of ln(FOODAWAY)", 
     xlab = "ln(FOODAWAY)", 
     col = "lightblue", 
     border = "black")

# 計算 ln(FOODAWAY) 的統計數據
mean_log_foodaway <- mean(cex5_small$log_foodaway, na.rm = TRUE)
median_log_foodaway <- median(cex5_small$log_foodaway, na.rm = TRUE)
percentile_25_log <- quantile(cex5_small$log_foodaway, 0.25, na.rm = TRUE)
percentile_75_log <- quantile(cex5_small$log_foodaway, 0.75, na.rm = TRUE)

# 顯示結果
cat("\nSummary statistics for ln(FOODAWAY):\n")
cat("Mean ln(FOODAWAY):", mean_log_foodaway, "\n")
cat("Median ln(FOODAWAY):", median_log_foodaway, "\n")
cat("25th Percentile ln(FOODAWAY):", percentile_25_log, "\n")
cat("75th Percentile ln(FOODAWAY):", percentile_75_log, "\n")

# 檢查 ln(FOODAWAY) 是否有 NA
cat("Number of missing ln(FOODAWAY) values:", sum(is.na(cex5_small$log_foodaway)), "\n")
```

## **(d) Quadratic Regression Curve with Tangent Line at 2000 Square Feet**

```{r}
# 確保 foodaway > 0，因為要取 log
cex5_small <- cex5_small[cex5_small$foodaway > 0, ]

# 取 ln(FOODAWAY)
cex5_small$log_foodaway <- log(cex5_small$foodaway)

# 執行線性回歸 ln(FOODAWAY) ~ INCOME
model <- lm(log_foodaway ~ income, data = cex5_small)

# 顯示回歸結果
summary(model)
```

## **(e) Scatter Plot of ln(FOODAWAY) vs. INCOME with Fitted Line**

```{r}
# 繪製散佈圖
plot(cex5_small$income, cex5_small$log_foodaway, 
     main = "Scatterplot of ln(FOODAWAY) vs. INCOME", 
     xlab = "INCOME (in $100 units)", 
     ylab = "ln(FOODAWAY)", 
     col = "blue", 
     pch = 16, 
     cex = 0.7)

# 加入回歸線
abline(model, col = "red", lwd = 2)
```

## **(f) Residual Plot from ln(FOODAWAY) Regression**

```{r}
# 計算殘差（實際值 - 預測值）
cex5_small$residuals <- residuals(model)

# 繪製殘差對 INCOME 的散佈圖
plot(cex5_small$income, cex5_small$residuals, 
     main = "Residuals vs. INCOME", 
     xlab = "INCOME (in $100 units)", 
     ylab = "Residuals", 
     col = "blue", 
     pch = 16, 
     cex = 0.7)

# 加入 y=0 的參考線
abline(h = 0, col = "red", lwd = 2)
```

# **CH2 Question 28**

## **(a) Summary Statistics and Histograms of WAGE and EDUC**

```{r}
# 1. 載入數據
load("cps5_small.RData")  # 確保 .RData 檔案名稱正確

# 2. 確認數據框的名稱
ls()  # 檢查載入的變數名稱
str(cps5_small)  # 確認數據結構，檢查 WAGE 和 EDUC 是否存在

# 3. 獲取 WAGE 和 EDUC 的摘要統計量
summary(cps5_small$wage)
summary(cps5_small$educ)

# 4. 計算平均值、中位數、標準差
mean_wage <- mean(cps5_small$wage, na.rm = TRUE)
median_wage <- median(cps5_small$wage, na.rm = TRUE)
sd_wage <- sd(cps5_small$wage, na.rm = TRUE)

mean_educ <- mean(cps5_small$educ, na.rm = TRUE)
median_educ <- median(cps5_small$educ, na.rm = TRUE)
sd_educ <- sd(cps5_small$educ, na.rm = TRUE)

# 5. 顯示統計結果
cat("WAGE Summary:\n")
cat("Mean:", mean_wage, " Median:", median_wage, " SD:", sd_wage, "\n\n")

cat("EDUC Summary:\n")
cat("Mean:", mean_educ, " Median:", median_educ, " SD:", sd_educ, "\n\n")

# 6. 繪製 WAGE 的直方圖
hist(cps5_small$wage, 
     main = "Histogram of Hourly Wage Rates", 
     xlab = "Hourly Wage", 
     col = "lightblue", 
     border = "black")

# 7. 繪製 EDUC 的直方圖
hist(cps5_small$educ, 
     main = "Histogram of Years of Education", 
     xlab = "Years of Education", 
     col = "lightgreen", 
     border = "black")
```

## **(b) Linear Regression of WAGE on EDUC**

```{r}
# 1. 執行線性回歸：WAGE ~ EDUC
model_wage <- lm(wage ~ educ, data = cps5_small)

# 2. 顯示回歸結果
summary(model_wage)
```

## **(c) Residual Analysis of the Linear Regression Model**

```{r}
# 1. 計算殘差（實際值 - 預測值）
cps5_small$residuals <- residuals(model_wage)

# 2. 繪製殘差對 EDUC 的散佈圖
plot(cps5_small$educ, cps5_small$residuals, 
     main = "Residuals vs. EDUC", 
     xlab = "Years of Education", 
     ylab = "Residuals", 
     col = "blue", 
     pch = 16, 
     cex = 0.7)

# 3. 加入 y = 0 的基準線
abline(h = 0, col = "red", lwd = 2)
```

## **(d) Separate Regressions by Gender and Race**

```{r}
# 1. 針對男性進行回歸
model_male <- lm(wage ~ educ, data = cps5_small, subset = (female == 0))
summary(model_male)

# 2. 針對女性進行回歸
model_female <- lm(wage ~ educ, data = cps5_small, subset = (female == 1))
summary(model_female)

# 3. 針對黑人 (black == 1) 進行回歸
model_black <- lm(wage ~ educ, data = cps5_small, subset = (black == 1))
summary(model_black)

# 4. 針對白人 (black == 0) 進行回歸
model_white <- lm(wage ~ educ, data = cps5_small, subset = (black == 0))
summary(model_white)
```

## **(e) Quadratic Regression of WAGE on EDUC**

```{r}
# 1. 執行只有 EDUC² 的回歸
model_quad_only <- lm(wage ~ I(educ^2), data = cps5_small)

# 2. 顯示回歸結果
summary(model_quad_only)

# 3. 計算邊際效果
coef_quad_only <- coef(model_quad_only)  # 提取係數
alpha2_only <- coef_quad_only["I(educ^2)"]

# 4. 邊際效果 (Marginal Effect) 計算
# d(WAGE)/d(EDUC) = 2 * alpha2 * EDUC
marginal_effect_12_only <- 2 * alpha2_only * 12
marginal_effect_16_only <- 2 * alpha2_only * 16

# 5. 顯示結果
cat("Marginal Effect at EDUC = 12:", marginal_effect_12_only, "\n")
cat("Marginal Effect at EDUC = 16:", marginal_effect_16_only, "\n")
```
## **(f) Comparison of Linear and Quadratic Regression Models**

```{r}
# 1. 生成擬合值 (Fitted Values)
cps5_small$fitted_linear <- fitted(model_wage)        # 線性回歸擬合值
cps5_small$fitted_quad_only <- fitted(model_quad_only)  # 只有 EDUC² 的回歸擬合值

# 2. 繪製散點圖 (原始數據)
plot(cps5_small$educ, cps5_small$wage, 
     main = "Comparison of Linear vs. Quadratic (EDUC²) Fit",
     xlab = "Years of Education", 
     ylab = "Hourly Wage", 
     col = "gray", 
     pch = 16, 
     cex = 0.7)

# 3. 繪製線性回歸擬合線
lines(sort(cps5_small$educ), sort(cps5_small$fitted_linear), 
      col = "blue", lwd = 2, lty = 2)  # 藍色虛線

# 4. 繪製只有 EDUC² 回歸擬合線
lines(sort(cps5_small$educ), sort(cps5_small$fitted_quad_only), 
      col = "red", lwd = 2)  # 紅色實線

# 5. 添加圖例
legend("topleft", legend = c("Linear Fit", "Quadratic (EDUC²) Fit"),
       col = c("blue", "red"), lty = c(2,1), lwd = 2)
```

